\chapter{Changelog}

\newcommand{\ChangeLogEntry}[3]{\subparagraph{Changes by #1 on #2} \begin{itemize} #3 \end{itemize} }

\tiny
\ChangeLogEntry{Tristan Lins}{08.09.2011}{
	\item Initial draft containing document description and package definition.
}
\ChangeLogEntry{Tristan Lins}{09.09.2011}{
	\item added PACKAGE XML information file.
	\item small adjustments in PACKAGE INI Format.
	\item elaborated usage of versions with an extension component.
}
\ChangeLogEntry{Tristan Lins}{10.09.2011}{
	\item Added first description of ER3 Repository, ER3 Repository Manager and ER3 Portal.
}
\ChangeLogEntry{Tristan Lins}{11.09.2011}{
	\item Removed INI Meta file specification.
	\item Reworked PACKAGE information file.
	\item Reworked XML schema.
	\item Reworked XML example.
}
\ChangeLogEntry{Tristan Lins}{12.09.2011}{
	\item Reworked XML Schema and example.
	\item added "versioning of dependencies" and "up/downgrade compatibility" to PACKAGE information.
	\item Removed description of license file and EULA.
}
\ChangeLogEntry{Christian Schiffler}{13.09.2011-14.09.2011}{
	\item Transition to \LaTeX{}, 
	\item translation to English and major rephrasing of all sections
	\item removed \techTerm{lowest\_bound} and \techTerm{highest\_bound} from \DisplFile{extension.xsd}.
}
\ChangeLogEntry{Tristan Lins}{15.09.2011}{
	\item Reworked Repository Portal\supref{sec:repository portal} section
}
\ChangeLogEntry{Christian Schiffler}{16.09.2011}{
	\item finished translation to English (mainly repository\supref{sec:repository} and repository manager \supref{sec:repository manager}).
}
\ChangeLogEntry{Tristan Lins}{19.09.2011}{
	\item Removed package scrhack, it generates the firstpage with "stylistings".
	\item Added first description of "Installation"\supref{sec:installation}, "Installation Process"\supref{sec:installation process} and "Installer"\supref{sec:installer interfaces and classes}.
}
\ChangeLogEntry{Tristan Lins}{25.09.2011}{
	\item Rework complete chapter "Installation"
	\item Added section "ER3 Manager"\supref{sec:er3 manager}.
	\item Added section "Contao Installation Manager"\supref{sec:contao installation manager}.
	\item Rename section "Installer" to "Installer interfaces and classes"\supref{sec:installer interfaces and classes}.
	\item Added subsection "InstallerInformation"\supref{sec:installerinformation}.
}
\ChangeLogEntry{Tristan Lins}{03.10.2011}{
	\item Added subsection "Extension Builder"\supref{sec:extension builder} to section "Repository Portal"\supref{sec:repository portal}
	\item Added section "Name spaces"\supref{sec:namespaces}.
}
\ChangeLogEntry{Tristan Lins}{12.10.2011}{
	\item Added subsection "Comments and rating"\supref{sec:comments and rating} to section "Repository Portal"\supref{sec:repository portal}
}
\ChangeLogEntry{Tristan Lins}{16.10.2011}{
	\item Added encrypted flag to "PACKAGE information file schema"\supref{sec:package information file schema}
	\item Added chapter Licensing
	\item Added appendix "Unencrypted package checksum calculation (upc.php)"\supref{sec:unencrypted package checksum calculation}
	\item Added appendix "Package encryption (encrypt.php)"\supref{sec:package encryption implementation}
	\item Added appendix "Installation id calculation (installationid.php)"\supref{sec:unencrypted package checksum calculation}
	\item Added example "Hello World!" extension package
}
\ChangeLogEntry{Tristan Lins}{25.10.2011}{
	\item Added chapter Certificates\supref{sec:certificates}
	\item Added chapter Extension store\supref{sec:extension store}
	\item Update external hyper links, now they works correctly.
}
\ChangeLogEntry{Tristan Lins}{26.10.2011}{
	\item Added chapter Missing topics\supref{sec:missing topics}
}
\ChangeLogEntry{Tristan Lins}{05.11.2011}{
	\item Added implementation of Version class
	\item Added unit test for Version class
}
\ChangeLogEntry{Tristan Lins}{13.11.2011}{
	\item Added chapter groups and extension management access
	\item Added chapter forks and successors
	\item Added chapter lifetime
}

\ChangeLogEntry{Christian Schiffler}{16.05.2012}{
	\item fixed typos
}
\normalsize