\label{sec:installation} 
\chapter{Installation}

This chapter describe how the Installation goes on and what programming basics are required.

\section[sec:er3 manager]{ER3 Manager}

The ER3 Manager (fka ER Client) is designed for Contao 2. It's a full replacement for the ER2 Client, that is currently in use.
It's a combined manager, that use extensions from the ER2 and the ER3. The ER3 Manager - or Extension Manager - have to manage all installed extension in an existing Contao 2 installation.

\section[sec:contao installation manager]{Contao Installation Manager}

The Contao Installation Manager (short CIM), is an Contao independend tool. It's designed for Contao 3 and should manage a complete Contao installation, including the installed Contao core and the installed extensions.
Summarized it's a combination of the current existing Install Tool and the ER3 Manager.

\paragraph{}
CIM should replace the Contao distribution package.
The idea behind CIM is, that a user only have to load a single PHP file, the CIM Installer.
The CIM Installer load all required CIM files.
Then the CIM provide a wizard to install a complete Contao.
Maybe with a basic template and basic extensions.

\paragraph{}
To make this possible, contao have to managed as a repository package itself.
This means, the CIM can manage the Contao core like any other extension.
In this case, there are no need for special behaviours to manage dependencies to the Contao Core.

\paragraph{}
But CIM is more than an installation tool.
CIM check the environment, if its compatible with Contao and provide solutions to make it compatible
For example activate the FTP file acces (fka SafeModeHack).
CIM manage the Contao system settings.
For example the database connection.
CIM simplified the installation process by automating the installation process.

\paragraph{}
For users, CIM is an absolute simplification of the installation process.
First download the CIM Installer (a small PHP file) and upload to the webspace.
Second call the CIM Installer, the CIM Installer download all required CIM files.
Third make shure, the environment is ready to install.
Fourth load a basic installation (Music Academy).
After CIM finish it work, the installation is ready to use.

\paragraph{}
For web workers/agencies additionaly it automating the installation process.
The automating process can be customized, to mind the custom preferences.
